### 2012-02-03 papers/itsc_2012/draft.tex
fixed exceeding speed limit: agents were trying TOO hard to follow others

\documentclass[12pt]{article}

\title{Paper}
\author{Dustin Carlino and Mike Depinet}
\date{...}

\begin{document}
\maketitle

% what are our goals with this paper?
% * demonstrate use of OSM and motivate more data collection
% * present a simulation framework
% * present a pretty naive static planning
% * motivate and demonstrate wards for hierarchial planning (but also reasoning
%    about congestion)

% contribution: autonomous car / human car sim on EXISTING INFRASTRUCTURE

\section{Map Construction from OSM}

each step

limitations

proposals to change OSM to fix these issues

integrating with what other sources for spawning traffic, placing POI?

\section{Simulation}

limitations of timesteps and small edges. in reality, more computational power
and more reactivity

justify arbitrary constants, like deaccel for cars

\end{document}