### 2012-02-04 papers/itsc_2012/draft.tex
re-enabled edge trimming with more restraints.. it's at least a bit better

\documentclass[12pt]{article}

\title{Paper}
\author{Dustin Carlino and Mike Depinet}
\date{...}

\begin{document}
\maketitle

% what are our goals with this paper?
% * demonstrate use of OSM and motivate more data collection
% * present a simulation framework
% * present a pretty naive static planning
% * motivate and demonstrate wards for hierarchial planning (but also reasoning
%    about congestion)

% contribution: autonomous car / human car sim on EXISTING INFRASTRUCTURE

\section{Map Construction from OSM}

each step

limitations

proposals to change OSM to fix these issues

integrating with what other sources for spawning traffic, placing POI?

\section{Simulation}

limitations of timesteps and small edges. in reality, more computational power
and more reactivity

justify arbitrary constants, like deaccel for cars

interface... what invariants do agents have to follow, and how does the default
behavior provide them?

\section{notes}

possible background/links
http://www.caliper.com/transmodeler/default.htm
http://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=6082793
 https://www.google.com/search?sourceid=chrome&ie=UTF-8&q=Agent-Based+Traffic+Simulation+and+Traffic+Signal+Timing+Optimization+with+GPU

\end{document}